Hire Ed: Job Market Dynamics for Tenure-Track Faculty Positions in Archaeology

### 2025-02-04 analysis/paper/article.qmd
add renv

Attend any workshop, conference, or panel that includes early career researchers and the conversation will be steered inexorably towards the academic job market: who is hiring, who has attained a tenure-track position, and who is out of luck this season. This obsessive focus on career trajectories is warranted. As of 2019, 53.5% of colleges and universities had replaced tenure-eligible lines with contingent positions [@the20222022]; today, 71% of faculty in the US are non-tenure-track [@culver] . The erosion of permanent jobs in American higher education is linked to a complex intersection of political and economic factors [@beck], including decreasing federal support for higher education, concomitant increases in student debt [@gusterson2017], a pronounced shift in university investment from faculty to administration [@graeber2018, pp. 162–163], and the growing privatization and market orientation of scientific research [@mirowski2011].

@@

>>>>

Tenure-track positions are a type of academic job that typically involve a probationary period (e.g. 5-7 years) culminating in an evaluation of research, teaching, and service, that if successful, results in permanent employment, job security and academic freedom. Attend any workshop, conference, or panel that includes early career researchers and the conversation will be steered inexorably towards the academic job market: who is hiring, who has attained a tenure-track position, and who is out of luck this season. This obsessive focus on career trajectories is warranted. As of 2019, 53.5% of colleges and universities had replaced tenure-eligible lines with contingent positions [@the20222022]; today, 71% of faculty in the US are non-tenure-track [@culver]. The erosion of permanent jobs in American higher education is linked to a complex intersection of political and economic factors [@beck], including decreasing federal support for higher education, concomitant increases in student debt [@gusterson2017], a pronounced shift in university investment from faculty to administration [@graeber2018, pp. 162–163], and the growing privatization and market orientation of scientific research [@mirowski2011].

@@

# Funding Statement

This research received no specific grant funding from any funding agency, commercial or not-for-profit sectors.

# Competing Interests Statement

Competing interests: The authors declare none

# Supplementary materials

Our supplementary materials are available at http://doi.org/xxx/xxx.
